OpReg-Boost: Learning to Accelerate Online Algorithms \\ with Operator Regression

### 2021-01-18 paper/main.tex
drafted simulations in paper

% new math commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\f}{\mathbold{f}}
\newcommand{\bdelta}{\mathbold{\delta}}
\newcommand{\transp}{\top}
\newcommand{\bv}{\mathbold{b}}
\newcommand{\e}{\mathbold{e}}
\newcommand{\p}{\mathbold{p}}
\newcommand{\q}{\mathbold{q}}
\newcommand{\tv}{\mathbold{t}}
\newcommand{\x}{\mathbold{x}}
\newcommand{\y}{\mathbold{y}}
\newcommand{\vv}{\mathbold{v}}
\newcommand{\z}{\mathbold{z}}
\newcommand{\xx}{\pmb{\xi}}
\newcommand{\ball}[2]{\mathbb{B}_{#2}(#1)}
\newcommand{\Am}{\mathbold{A}}
\renewcommand{\Im}{\mathbold{I}}
\newcommand{\Pm}{\mathbold{P}}
\newcommand{\cv}{\mathbold{c}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}

@@

\andrea{Decide $k$ or $k+1$}

@@

@@

\begin{itemize}
\item non-negative Least-squares

\item L1 ?

@@

\item weakly convex: time-varying phase retrival? (see Convergence of a Stochastic Gradient Method with Momentum for Nonsmooth Nonconvex Optimization)

@@

%------------------------------------------------
\subsection{Simulations set-up}
We consider the following time-varying problem:
\begin{equation}\label{eq:tv-problem}
	\x^*(t_k) = \argmin_{\x \in \R^n} \frac{1}{2} \norm{\Am \x - \bv(t_k)}^2 + w \norm{\x}_1
\end{equation}
with $n = 10$, $\Am$ matrix with maximum and minimum (non-zero) eigenvalues $\sqrt{L} = 10^8$, $\sqrt{\mu} = 1$, and with rank $5$; $\y(t_k)$ has sinusoidal components with $3$ zero components. Due to $\Am$ being rank deficient, the cost $f$ is convex but not strongly so.

%------------------------------------------------
\subsection{Results}

%% FIGURE
%\begin{figure}[ht]
%\vskip 0.2in
%\begin{center}
%\centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
%\caption{Historical locations and number of accepted papers for International
%Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
%Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
%produced, the number of accepted papers for ICML 2008 was unknown and instead
%estimated.}
%\label{icml-historical}
%\end{center}
%\vskip -0.2in
%\end{figure}

%% ALGORITHM
%\begin{algorithm}[tb]
%   \caption{Bubble Sort}
%   \label{alg:example}
%\begin{algorithmic}
%   \STATE {\bfseries Input:} data $x_i$, size $m$
%   \REPEAT
%   \STATE Initialize $noChange = true$.
%   \FOR{$i=1$ {\bfseries to} $m-1$}
%   \IF{$x_i > x_{i+1}$}
%   \STATE Swap $x_i$ and $x_{i+1}$
%   \STATE $noChange = false$
%   \ENDIF
%   \ENDFOR
%   \UNTIL{$noChange$ is $true$}
%\end{algorithmic}
%\end{algorithm}

%% TABLE
%\begin{table}[t]
%\caption{Classification accuracies for naive Bayes and flexible
%Bayes on various data sets.}
%\label{sample-table}
%\vskip 0.15in
%\begin{center}
%\begin{small}
%\begin{sc}
%\begin{tabular}{lcccr}
%\toprule
%Data set & Naive & Flexible & Better? \\
%\midrule
%Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
%Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
%Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
%Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
%Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
%Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
%Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
%Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
%\bottomrule
%\end{tabular}
%\end{sc}
%\end{small}
%\end{center}
%\vskip -0.1in
%\end{table}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width = 0.9\columnwidth]{Figures/comparison_projection_methods}
\caption{Comparison of feasibility and best approximation methods. \nicola{modified the figure to have markers}}
\label{fig:projection_methods}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width = 0.9\columnwidth]{Figures/MAP_histogram}
\caption{Histogram of iterations required by MAP over $10000$ repetitions.}
\label{fig:map-histogram}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width = 0.9\columnwidth]{Figures/MAP_changing_d}
\caption{Histogram of iterations required by MAP over $10000$ repetitions.}
\label{fig:map-changing-d}
\end{figure}

>>>>

% new math commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\f}{\mathbold{f}}
\newcommand{\bdelta}{\mathbold{\delta}}
\newcommand{\transp}{\top}
\newcommand{\av}{\mathbold{a}}
\newcommand{\bv}{\mathbold{b}}
\newcommand{\e}{\mathbold{e}}
\newcommand{\p}{\mathbold{p}}
\newcommand{\q}{\mathbold{q}}
\newcommand{\tv}{\mathbold{t}}
\newcommand{\x}{\mathbold{x}}
\newcommand{\y}{\mathbold{y}}
\newcommand{\vv}{\mathbold{v}}
\newcommand{\z}{\mathbold{z}}
\newcommand{\xx}{\pmb{\xi}}
\newcommand{\ball}[2]{\mathbb{B}_{#2}(#1)}
\newcommand{\Am}{\mathbold{A}}
\renewcommand{\Im}{\mathbold{I}}
\newcommand{\Pm}{\mathbold{P}}
\newcommand{\cv}{\mathbold{c}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}

@@

\andrea{Decide $k$ or $k+1$} \nicola{we could go with $\x_{k+1}$ coming from $f_k$, which is what the learning community usually does}

@@

\nicola{say that the algorithm applies to any operator not only forward-backward}

@@

\begin{itemize}
\item non-negative Least-squares \nicola{not very interesting, the results are similar to lasso}

\item L1 \nicola{done}

@@

\item weakly convex: time-varying phase retrival? (see Convergence of a Stochastic Gradient Method with Momentum for Nonsmooth Nonconvex Optimization) \nicola{done}

@@

In the following we present a series of experiments to evaluate the performance of the proposed methods. We consider the following challenging set-ups: \emph{(i)} online linear regression with a cost that has an extremely high condition number ($10^8$) and is convex but not strongly so; \emph{(ii)} an online phase retrieval problem that is weakly convex, and which is characterized by a high computational cost for an operator evaluation, \nicola{\emph{(iii)} ....}

%------------------------------------------------
\subsection{Online linear regression}
In this first experiment we consider the following time-varying problem:
\begin{equation}\label{eq:linear-regression}
	\x^*(t_k) \in \argmin_{\x \in \R^n} \frac{1}{2} \norm{\Am \x - \bv(t_k)}^2 + w \norm{\x}_1
\end{equation}
with $n = 10$, $w = 5$, $\Am$ matrix with maximum and minimum (non-zero) eigenvalues $\sqrt{L} = 10^4$, $\sqrt{\mu} = 1$, and with rank $5$; $\y(t_k)$ has sinusoidal components with $3$ zero components. Due to $\Am$ being rank deficient, the cost $f$ is convex but not strongly so, and the condition number of the regression cost is $\kappa = L / \mu = 10^8$.

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{Figures/lasso-operator_calls}}
\caption{Comparison of the cumulative tracking error evolution for forward-backward, its accelerated versions FISTA \cite{beck_fast_2009} and Anderson \cite{pmlr-v119-mai20a}, OpReg-Boost and interpolated OpReg-Boost, and CvxReg-Boost. The methods are given a budget of $10$ gradient calls per sampling time.}
\label{fig:lasso-operator_calls}
\end{center}
\vskip -0.2in
\end{figure}

%------------------------------------------------
\subsection{Online phase retrieval}
In this experiment we consider the following online phase retrieval problem \cite{duchi_stochastic_2018}
\begin{equation}\label{eq:phase-retrieval}
	\x^*(t_k) \in \argmin_{\x \in \R^n} \frac{1}{m} \sum_{i = 1}^{m} \left| \langle \av_i, \x \rangle^2 - b_i(t_k) \right|
\end{equation}
where we want to reconstruct a time-varying signal $\y(t_k)$ (which is performing a random walk on the unit sphere in $\R^n$) from the noisy measurements $b_i(t_k) = \langle \av_i, \y(t_k) \rangle^2 + \xi_i(t_k)$. In particular, we have $n = 15$ and $m = 200$, the measurement noises $\xi_i(t_k)$ are drawn from a Laplacian distribution of scale $1$, and $\Am = \begin{bmatrix} \av_1 & \cdots & \av_m \end{bmatrix}^\top = \mathbold{U} \mathbold{D}$ where $\mathbold{U}$ is a randomly generated orthogonal matrix and $\mathbold{D}$ is a diagonal matrix with elements equispaced in $[0, 100]$.

We consider the prox-linear solver proposed in \cite{drusvyatskiy_error_2018} and characterized by $\T_k \y = \prox_{\alpha f_\y(\cdot; t_k)}(\y)$ where we $f_\y$ denotes the linearized version of the cost in~\eqref{eq:phase-retrieval}:
$$
	f_\y(\x; t_k) = \frac{1}{m} \sum_{i = 1}^m \left| \langle \av_i, \y \rangle^2 + 2 \langle \av_i, \y \rangle \langle \av_i, \x - \y \rangle - b_i \right|.
$$
Since we do not have a closed form expression for $\T_k$, each operator evaluation requires the solution of a quadratic program.

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{Figures/phase_retrieval}}
\caption{Comparison of the cumulative tracking error evolution for prox-linear \cite{drusvyatskiy_error_2018}, the Anderson acceleration version, and OpReg-Boost. The methods are given a budget of $10$ operator evaluations per sampling time.}
\label{fig:phase-retrieval}
\end{center}
\vskip -0.2in
\end{figure}

@@

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{Figures/comparison_projection_methods}}
\caption{Comparison of feasibility and best approximation methods.}
\label{fig:projection_methods}
\end{center}
\vskip -0.2in
\end{figure}

@@

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{Figures/MAP_histogram}}
\caption{Histogram of iterations required by MAP over $10000$ repetitions.}
\label{fig:map-histogram}
\end{center}
\vskip -0.2in
\end{figure}

@@

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{Figures/MAP_changing_d}}
\caption{Mean number of iterations required by MAP as a function of the number of sets, computed over $10000$ repetitions. The dashed curve depicts a logarithmic interpolation of the curve.}
\label{fig:map-changing-d}
\end{center}
\vskip -0.2in
\end{figure}
